---
title: "Simulation Applications - Bonus HW5"
author: "Cameron R. Shabahang"
---

```{r}
#Robust random number generator 
MYUNIFORM <- function(inputvar) {
# Initialize Constants
  IM1<-2147483563
  IM2<-2147483399
  IA1<-40014
  IA2<-40692
  IQ1<-53668
  IQ2<-52774
  IR1<-12211
  IR2<-3791
  NTAB<-32
  EPS<-1.2e-7
  RNMX<-1.-EPS
# Transform Variables
  IMM1<-IM1-1
  NDIV<-as.integer(1+IMM1/NTAB)
  AM<-1.0/IM1
# Initialize variables and arrays
  idum<-inputvar[1]
  idum2<-123456789
  numran<-inputvar[2]
  ran2<-0
  iy<-0
  iv<-rep(0,NTAB)
  rand_uniform_c<-rep(0,numran)
# Run the random number loop
  icount<-1
  for (icount in 1:numran) {
    if (idum <= 0) {
      idum<-max(-idum,1)
      idum2<-idum
      j<-NTAB+8
      while (j > 0) {
        k=as.integer(idum/IQ1)
        idum<-IA1*(idum-k*IQ1)-k*IR1
        if (idum < 0) {idum=idum+IM1}
        if (j <= NTAB) {iv[j]<-idum}
        j<-j-1
      }
        iy<-iv[1]
      }
    k<-as.integer(idum/IQ1)
    idum<-IA1*(idum-k*IQ1)-k*IR1
    if(idum < 0) {idum=idum+IM1}
    k=as.integer(idum2/IQ2)
    idum2<-IA2*(idum2-k*IQ2)-k*IR2 
    if (idum2 < 0) {idum2<-idum2+IM2}
    j<-as.integer(iy/NDIV)+1
    iy<-iv[j]-idum2
    iv[j]<-idum
    if(iy < 1) {iy<-iy+IMM1}
    ran2<-min(AM*iy,RNMX)
    rand_uniform_c[icount]<-ran2
  }
  return(rand_uniform_c)
}

MYNORM <- function(seed, x) {
  
  rand_c <- MYUNIFORM(c(seed,x))
# Initialize Constants
  a0<-2.50662823884
  a1<--18.61500062529
  a2<-41.39119773534
  a3<--25.44106049637
  b0<--8.47351093090
  b1<-23.08336743743
  b2<--21.06224101826
  b3<-3.13082909833
  c0<-0.3374754822726147
  c1<-0.9761690190917186
  c2<-0.1607979714918209
  c3<-0.0276438810333863
  c4<-0.0038405729373609
  c5<-0.0003951896511919
  c6<-0.0000321767881768
  c7<-0.0000002888167364
  c8<-0.0000003960315187
# Loop over set of uniform random numbers and transform
  jcount<-1
  numran<-x
  rand_norm_c<-rep(0,numran)
  while(jcount <= numran) {
    u<-rand_c[jcount]
    y<-u-0.5
    if(abs(y) < 0.42) {
      r<-y*y
      x<-y*(((a3*r+a2)*r+a1)*r+a0)/((((b3*r+b2)*r+b1)*r+b0)*r+1)
    } else {
      r<-u
      if(y>0){r<-1-u}
      r<-log(-log(r))
      x<-c0+r*(c1+r*(c2+r*(c3+r*(c4+r*(c5+r*(c6+r*(c7+r*c8)))))))
      if(y<0){x<--x}
    }
  rand_norm_c[jcount]<-x
  jcount=jcount+1
  }
  return(rand_norm_c)
}
```
Calculate the 1% VaR level for a bond portfolio consisting of 100 bonds through Monte Carlo simulation.  Simulation parameters are as follows:

(1)	Populate a portfolio of 100 Bonds using a uniform random number generator
(2)	Assume that the probability of default is determined through a multivariate normal process.  If a normally distributed random number is below one standard deviation from a mean of zero, the bond is in default.
(3)	Simulate 100 random numbers from a multivariate normal distribution.  Randomly populate a 100 x 100 matrix to represent the covariance matrix of the multivariate normal. 
(4)	Using the covariance matrix generate 100 random numbers representing changes in the value of the asset in one day.  Check if any of the random numbers is less than one standard deviation from zero to signal default.
(5)	Assume that the bond loses 20% of its value if there is default.  Calculate the level of loss of the portfolio.
(6)	Repeat the steps above to generate a probability distribution for losses.  Determine the level and percentage loss  of losses at the 5% probability level.


```{r}
seed<--1000
numreg<-100

var1<-rep(0,numreg)
pctloss1<-rep(0,numreg)
var5<-rep(0,numreg)
pctloss5<-rep(0,numreg)

s1v<-rep(0,numreg)
s2v<-rep(0,numreg)
s3v<-rep(0,numreg)
s4v<-rep(0,numreg)
s5v<-rep(0,numreg)

v1v<-rep(0,numreg)
v2v<-rep(0,numreg)
v3v<-rep(0,numreg)
v4v<-rep(0,numreg)
v5v<-rep(0,numreg)

sv1v<-rep(0,numreg)
sv2v<-rep(0,numreg)
sv3v<-rep(0,numreg)
sv4v<-rep(0,numreg)
sv5v<-rep(0,numreg)

c1v<-rep(0,numreg)
c2v<-rep(0,numreg)
c3v<-rep(0,numreg)
c4v<-rep(0,numreg)
c5v<-rep(0,numreg)

cm1v<-rep(0,numreg)
cm2v<-rep(0,numreg)
cm3v<-rep(0,numreg)
cm4v<-rep(0,numreg)
cm5v<-rep(0,numreg)

vavg<-rep(0,numreg)
cavg<-rep(0,numreg)

#Populate a portfolio of a hundred bonds using a random number generator 
for (rcount in 1:numreg) {
  DNum<-100
  inputvar<-c(seed,DNum)
  rand_uniform_c<-MYUNIFORM(inputvar)
  ran100<-as.integer(rand_uniform_c*100)+1
  
  portfolio<-rep(0,DNum)
  portfolio2<-rep(0,DNum)
  for (icount in 1:DNum) {
    portfolio[icount]<-ran100[icount]*10000
  }
  seed<--ran100[icount]
  if(rcount == 1) cat(sum(portfolio),"\n")
  if(rcount == 1) cat(portfolio,"\n")

  numran<-200000
  seed2<--1000 #Shocks to asset values
  rand_norm_c<-MYNORM(seed2,numran)
  
  
  AMatrix <- matrix(rnorm(rand_norm_c^2), DNum, DNum) #
  CovarM <- (crossprod(AMatrix))/100 #Covariance matrix via cross product 
  AMatrix<-chol(CovarM) #Cholesky decomposition
  AMatrixt<-t(AMatrix) #Transposition of the matrix

  DeltaT<-1/365 #Timestep
  Volatility<-diag(CovarM)
  Volatility<-sqrt(Volatility)
  lcount<-1
  randZ<-rep(0,DNum)
  numruns<-1000
  P0<-sum(portfolio)
  rets<-rep(0,numruns)
  dollar_rets<-rep(0,numruns)
  for (icount in 1:numruns) {
    for (jcount in 1:DNum) {
      randZ[jcount]<-rand_norm_c[lcount]
      lcount<-lcount+1
    }
    ZVals<-AMatrixt%*%randZ
    sdz<-sd(randZ)
    for (jcount in 1:DNum) {
      portfolio2[jcount]<-portfolio[jcount]*(exp(((-0.5*Volatility[jcount]^2)*DeltaT)+(sqrt(DeltaT))*ZVals[jcount]))
      #check for loss greater than 1 std dev for default signal
      if(randZ[jcount]<(-sdz)){
        #20% haricut on bond if default signal
        portfolio2[jcount]<-.8*portfolio2[jcount]
      }
    }
    P1<-sum(portfolio2)
    rets[icount]<-log(P1/P0)
    dollar_rets[icount]<-P1-P0
  }
  if (rcount == 1) {hist(rets)}
  if (rcount == 1) {hist(dollar_rets,breaks=20)}
  quantile(dollar_rets,c(0.01,0.05,0.5,0.95,0.99))
  
  #
  # Regress for 1st Percentile
  #
  # Use Asset Size as a variable
  prank<-rank(portfolio2,ties.method=c("random"))
  for (icount in 1:DNum) {
    if(prank[icount]== 100) {s1 <- portfolio[icount]}
    if(prank[icount]== 99) {s2 <- portfolio[icount]}
    if(prank[icount]== 98) {s3 <- portfolio[icount]}
    if(prank[icount]== 97) {s4 <- portfolio[icount]}
    if(prank[icount]== 96) {s5 <- portfolio[icount]}
  }
  # vol of asset w biggest size (not highest vol)
  for (icount in 1:DNum) {
    if(prank[icount]== 100) {sv1 <- Volatility[icount]}
    if(prank[icount]== 99) {sv2 <- Volatility[icount]}
    if(prank[icount]== 98) {sv3 <- Volatility[icount]}
    if(prank[icount]== 97) {sv4 <- Volatility[icount]}
    if(prank[icount]== 96) {sv5 <- Volatility[icount]}
  }
  #
  #
  # Use Volatility as variable
  #
  # vol of asset w highest vol
  prank<-rank(Volatility,ties.method=c("random"))
  for (icount in 1:DNum) {
    if(prank[icount]== 100) {v1 <- Volatility[icount]}
    if(prank[icount]== 99) {v2 <- Volatility[icount]}
    if(prank[icount]== 98) {v3 <- Volatility[icount]}
    if(prank[icount]== 97) {v4 <- Volatility[icount]}
    if(prank[icount]== 96) {v5 <- Volatility[icount]}
  }
  #
  c1<-0
  c2<-0
  c3<-0
  c4<-0
  c5<-0
  CovarM1<-CovarM
  for (i in 1:DNum) {
        CovarM1[i,i]<--100 
  }
  #ignores variances - diagonal, only covariances
  c1<-max(CovarM1)
  
  #now set highest to -100 iteratively to get top5 cov
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==c1) {
        CovarM1[i,j]<--100 
        i<-DNum
        j<-DNum}
      }}
  c2<-max(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==c2) {
        CovarM1[i,j]<--100 
        i<-DNum
        j<-DNum}
      }}
  c3<-max(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==c3) {
        CovarM1[i,j]<--100 
        i<-DNum
        j<-DNum}
      }}
  c4<-max(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==c4) {
        CovarM1[i,j]<--100 
        i<-DNum
        j<-DNum}
      }}
  c5<-max(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==c5) {
        CovarM1[i,j]<--100 
        i<-DNum
        j<-DNum}
      }}
  #
  #same process for lowest cov's
  cm1<-0
  cm2<-0
  cm3<-0
  cm4<-0
  cm5<-0
  CovarM1<-CovarM
  cm1<-min(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==cm1) {
        CovarM1[i,j]<-+100 
        i<-DNum
        j<-DNum}
      }}
  cm2<-min(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==cm2) {
        CovarM1[i,j]<-+100 
        i<-DNum
        j<-DNum}
      }}
  cm3<-min(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==cm3) {
        CovarM1[i,j]<-+100 
        i<-DNum
        j<-DNum}
      }}
  cm4<-min(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==cm4) {
        CovarM1[i,j]<-+100 
        i<-DNum
        j<-DNum}
      }}
  cm5<-min(CovarM1)
  for (i in 1:DNum) {
    for (j in 1:DNum) {
      if( CovarM1[i,j]==cm5) {
        CovarM1[i,j]<-+100 
        i<-DNum
        j<-DNum}
      }}
  #
  var1[rcount]<-quantile(dollar_rets,0.01)
  pctloss1[rcount]<-(P0+var1[rcount])/P0-1
  var5[rcount]<-quantile(dollar_rets,0.05)
  pctloss5[rcount]<-(P0+var5[rcount])/P0-1
  s1v[rcount]<-s1
  s2v[rcount]<-s2
  s3v[rcount]<-s3
  s4v[rcount]<-s4
  s5v[rcount]<-s5
  v1v[rcount]<-v1
  v2v[rcount]<-v2
  v3v[rcount]<-v3
  v4v[rcount]<-v4
  v5v[rcount]<-v5
  sv1v[rcount]<-sv1
  sv2v[rcount]<-sv2
  sv3v[rcount]<-sv3
  sv4v[rcount]<-sv4
  sv5v[rcount]<-sv5
  c1v[rcount]<-c1
  c2v[rcount]<-c2
  c3v[rcount]<-c3
  c4v[rcount]<-c4
  c5v[rcount]<-c5
  cm1v[rcount]<-cm1
  cm2v[rcount]<-cm2
  cm3v[rcount]<-cm3
  cm4v[rcount]<-cm4
  cm5v[rcount]<-cm5
  vavg[rcount]<-mean(Volatility)
  cavg[rcount]<-mean(CovarM)-mean(Volatility)
}
```


```{r}
var_data = data.frame (
var1, var5, s1v, s2v, s3v, s4v, s5v, v1v, v2v, v3v, v4v, v5v, sv1v, sv2v, sv3v, sv4v, sv5v, c1v, c2v, c3v, c4v, c5v, cm1v, cm2v, cm3v, cm4v, cm5v, cavg, vavg
)

fit<-lm(var5 ~ s1v+ s2v+ s3v+ s4v+ s5v+ v1v+ v2v+ v3v+ v4v+ v5v+ sv1v+ sv2v+ sv3v+ sv4v+ sv5v+ c1v+ c2v+ c3v+ c4v+ c5v+ cm1v+ cm2v+ cm3v+ cm4v+ cm5v, data=var_data)
summary(fit)

fit1<-lm(var5 ~ s1v+ v1v+ c1v+ cm1v, data=var_data)
summary(fit1)

fit2<-lm(var5 ~ vavg + cavg, data=var_data)
summary(fit2)

```

We are conducting a 1% VaR analysis, so it is expected that var5 regressed against var1 will lead to an insignificant result. 

Regressing against the total average variance and average covariance, both factors are significant for a sample run. Other runs may not yield the same results. Once a threshold is reached, there can be large drops in portfolio value. 

```{r}
#Average 1% var across all simulations
cat("\nThe average 1% VaR over", numreg, "runs was", mean(var1), "for a daily loss of", -100*mean(pctloss1),"%")

#$ amount and % loss with 5% probability
cat("\n\nThere is a 5% probability of maximum losses reaching $", -mean(var5), "for a daily loss of return of", -mean(pctloss5)*100,"%")

```